Reviving stochasticity: eccentricity convergence depends on black hole binary orbit

### 2023-05-15 papers/paper-eccentricity/paper-eccentricity.tex
Add some ideas and drafting to the paper
Start writing about the toy model
- add to results section

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols

@@

% Please keep new commands to a minimum, and use \newcommand not \def to avoid
% overwriting existing commands. Example:
%\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared
\newcommand{\ketju}{\textsc{Ketju}}
\newcommand{\mstar}{\textsc{Mstar}}
\newcommand{\gadget}{\textsc{Gadget-3}}

\newcommand{\Msun}{\ensuremath{M_{\sun}}}
\newcommand{\Rvir}{\ensuremath{R_\mathrm{vir}}}

\newcommand{\dd}[1]{\ensuremath{\mathrm{d}#1}}
\newcommand{\dv}[2]{\ensuremath{\frac{\dd{#1}}{\dd{#2}}}}

@@

\textit{Software:} \ketju{} \citep{rantala17}, \gadget{} \citep{springel05}, \textsc{Stan} \citep{carpenter17}, NumPy \citep{numpy}, SciPy \citep{scipy}, Matplotlib \citep{matplotlib}, pygad \citep{rottgers20}, Arviz \citep{arviz}.

>>>>

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
%\usepackage{amssymb}	% Extra maths symbols

@@

% Please keep new commands to a minimum, and use \newcommand not \def to avoid
% overwriting existing commands. Example:
%\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared
\newcommand{\ketju}{\textsc{Ketju}}
\newcommand{\mstar}{\textsc{Mstar}}
\newcommand{\gadget}{\textsc{Gadget-4}}

\newcommand{\Msun}{\ensuremath{\mathrm{M}_{\sun}}}
\newcommand{\Rvir}{\ensuremath{R_\mathrm{vir}}}

\newcommand{\dd}[1]{\ensuremath{\mathrm{d}#1}}
\newcommand{\dv}[2]{\ensuremath{\frac{\dd{#1}}{\dd{#2}}}}
\newcommand{\vb}[1]{\ensuremath{\mathbf{#1}}}

@@

\textit{Software:} \ketju{} \citep{rantala17}, \gadget{} \citep{springel22}, \textsc{Stan} \citep{carpenter17}, NumPy \citep{numpy}, SciPy \citep{scipy}, Matplotlib \citep{matplotlib}, pygad \citep{rottgers20}, Arviz \citep{arviz}.

### 2023-05-15 papers/paper-eccentricity/sections/numerical.tex
- add to results section

\subsection{Merger ICs}
We construct isolated merger initial conditions by placing two galaxies on an elliptical orbit with varying values of eccentricity, at a fixed initial separation of $D=3.72\,\mathrm{kpc}$ and semimajor axis $a_0=2.79\,\mathrm{kpc}$.
We follow the same orbital configurations as \citet{gualandris22}, with one additional orbit. 
The chosen eccentricities $e$ are 0.5, 0.7, 0.9, and 0.99.
We ensure that the radial and tangential velocities of the initial merger setup are consistent with the values reported in \citet{gualandris22}.
An example of each orbital configuration is shown in \autoref{fig:orbits}.
For each orbital configuration, we run \drafting{X} realisations, to account for Brownian motion effects of the SMBH binary caused by the discretised phase space. 
Each orbital geometry is simulated using our fiducial particle resolution of $M_\bullet/m_\star = 10^4$.
Additionally, we run the $e=0.9$ and $e=0.99$ orbits at varying mass resolutions, using the galaxy ICs at the mass resolutions described in \autoref{ssec:gics}.

>>>>

\subsection{Merger ICs}
We construct isolated merger initial conditions by placing two galaxies on an elliptical orbit with varying values of eccentricity, at a fixed initial separation of $D=3.72\,\mathrm{kpc}$ and semimajor axis $a_0=2.79\,\mathrm{kpc}$.
We follow the same orbital configurations as \citet{gualandris22}, with three additional orbit. 
The chosen eccentricities $e$ are 0.5, 0.7, 0.9, 0.95, 0.97, and 0.99.
We ensure that the radial and tangential velocities of the initial merger setup are consistent with the values reported in \citet{gualandris22}.
An example of each orbital configuration is shown in \autoref{fig:orbits}.
For each orbital configuration, we run ten realisations, to account for Brownian motion effects of the SMBH binary caused by the discretised phase space. 
Each orbital geometry is simulated using our fiducial particle resolution of $M_\bullet/m_\star = 10^4$.
Additionally, we run the $e=0.9$ and $e=0.99$ orbits at varying mass resolutions, using the galaxy ICs at the mass resolutions described in \autoref{ssec:gics}.

### 2023-05-15 papers/paper-eccentricity/sections/results.tex
fix bug in KeplerModels that resulted in an extra group being added if a data extraction process failed
- add to results section

\subsection{Eccentricity scatter in low-e orbits}
\begin{itemize}
    \item Show that impact parameters not sensitive
\end{itemize}

\subsection{Eccentricity scatter in high-e orbits}
\begin{itemize}
    \item show that impact parameters sensitive
\end{itemize}
\begin{figure*}
    \includegraphics{nasim_scatter_e0-990.pdf}
    \caption{\drafting{Show a plot like this where the scatter in eccentricity is determined as the standard deviation across simulations}}
    \label{fig:nasim_e90}
\end{figure*}
We show the dependence of eccentricity standard deviation on mass resolution for the $e_0=0.99$ orbit in \autoref{fig:nasim_e90}. 
We find that the scaling of $\sigma_e$ is better described by a relation proportional to $1/N_\star$, as opposed to the $1/\sqrt{N_\star}$ relation described in \citet{nasim20}.
We attribute this dramatic decrease in $\sigma_e$ to an inflated value of $\sigma_e$ at low mass resolution: there is increased scatter in eccentricity at low mass resolutions for very radial initial orbits of the SMBH binary.

@@

\subsection{Explaining the eccentricity scatter with a toy model}

>>>>

\subsection{Eccentricity scatter in low-e orbits}
\begin{itemize}
    \item Show that impact parameters not sensitive
\end{itemize}
\begin{figure*}
    \includegraphics{nasim_scatter_e0-900.pdf}
    \caption{\drafting{Inverse modelling convergence of eccentricity scatter as a function of mass resolution for $e_0=0.90$. NOTE AXIS LABEL SHOULD BE MASS RESOLUTION NOT PARTICLE NUMBER HERE}}
    \label{fig:nasim_e90}
\end{figure*}
We show the dependence of eccentricity standard deviation on mass resolution for the $e_0=0.90$ orbit in \autoref{fig:nasim_e99}. 
Immediately apparent is the lack of convergence of $\sigma_e$, with $5\times10^3$ and $1\times10^4$ mass resolutions demonstrating an increased scatter compared to the lowest mass resolution of $2.5\times10^3$.
The peculiar values of $\sigma_e$ arises due to the assumed Gaussian distribution of the eccentricity between simulations.
Inspecting the orbital parameters for all simulations that form the $e_0=0.90$ suite, it is found that for resolutions of $5\times10^3$ and $1\times10^4$ the eccentricity clusters around a value of $e_\mathrm{h}\simeq0.90$, except for one simulation at each resolution which has a value of $e_\mathrm{h}<0.4$.
This one outlier for each of the two resolutions artificially inflates the measured eccentricity variance, and by removing these two points we recover the $1/\sqrt{N_\star}$ scaling of $\sigma_e$ as in \citet{nasim20}.
However, these low eccentricity values are a valid measurement from the simulation, and highlight the weakness in the inverse modelling approach to determine the spread in eccentricity. 
The cause of these low eccentricity values is discussed in \autoref{ssec:toy}.
\drafting{Run Bayesian model, more robust?}

\subsection{Eccentricity scatter in high-e orbits}
\begin{itemize}
    \item show that impact parameters sensitive
\end{itemize}
\begin{figure*}
    \includegraphics{nasim_scatter_e0-990.pdf}
    \caption{\drafting{Show a plot like this where the scatter in eccentricity is determined as the standard deviation across simulations}}
    \label{fig:nasim_e99}
\end{figure*}
We show the dependence of eccentricity standard deviation on mass resolution for the $e_0=0.99$ orbit in \autoref{fig:nasim_e99}. 
We find that the scaling of $\sigma_e$ is better described by a relation proportional to $1/N_\star$, as opposed to the $1/\sqrt{N_\star}$ relation described in \citet{nasim20}.
We attribute this dramatic decrease in $\sigma_e$ to an inflated value of $\sigma_e$ at low mass resolution: there is increased scatter in eccentricity at low mass resolutions for very radial initial orbits of the SMBH binary. 
\drafting{TODO: figure comparing different eccentricity scatters}

@@

\subsection{Explaining the eccentricity scatter with a toy model}\label{ssec:toy}
